Title

### 2022-11-17 Manuscript/Manuscript.Rmd
Quick comments from Nacho

1. Understanding the mechanisms driving community stability and whether this stability should be related to the stability of ecosystem functions is important for predicting the resilience of communities to changes in biodiversity. While community stability can be influenced by species richness and asynchrony of species,  the relation with the stability of ecosystem functioning provided by this community is less explored, and specially for functions involving several trophic levels such as plant pollination.

2. Using a 5-year dataset across 13 Mediterranean woodland patches, we assessed the effect of the richness and asynchrony of pollinators as drivers of the stability of plant's visitation rate and its relationship between the stability of visitation rate and the stability of the plants reproductive success. [In addition, we analyzed whether the year together with richness and visitation rate affected the proportion of fruits and the number of seeds.](esto pensamos como ponerlo con los results finales)

3. The stability of plant's visitation rate is more affected by the asynchrony of pollinator species abundances than by pollinator richness shows a negative tendency. As for the stability of the plant reproductive success, we found no relationship between the stability of fruits and the stability of the visitation rate, however a negative relation was observed between stability of seed and the stability of visitation rate. However, when years are analysed we found that in some years/[bad years](), the effect of pollinator richness and visitation rate on fruit proportion becomes important.

@@

The role of biodiversity in ecosystem function has long been of great interest to researchers (Hooper et al 2005, Gagic et al., 2015, Lefcheck et al 2015, Oliver et al., 2015, Hong et al 2021). We have experimental and observational evidence that biodiversity enhances ecosystem functioning (Cardinale et al., 2006, Cardinale et al., 2012, Tilman et al., 2014), but the role of biodiversity for the stability of ecosystem functioning is less explored (but see Griffin et al., 2009, Grman et al., 2010), specially for functions involving several trophic levels such as plant pollination (Ebeling et al., 2008, Winfree and Kremen 2009, Garibaldi et al 2011, [Lázaro et al 2022](wild Olea europaea communities)). Globally, it is estimated that at least 78% of wild plants and 75% of leading crops depend to some extent on pollination by animals (Ollerton et al., 2011; Kleijn et al., 2015). Understanding the mechanisms driving pollinator community stability and how this relates to the stability of ecosystem functions and services is important for predicting changes in both community structure and ecosystem function.

Stability can be defined as the inverse of the coefficient of variation of ecosystem properties over time (McCann, 2000; Tilman, 1999). Among main driving of stability are the diversity and species asynchrony. A higher number of species might provide a wider range of responses to environmental changes, which would result in a higher stability (McCann, 2000, Tilman et al. 2006, Hector et al. 2010, Xu et al. 2021), although non-significant relationships between diversity and stability have also been observed (Hooper et al. 2005, van der Plas 2019). The stability is also influenced by the degree of asynchrony between species. In this sense, higher levels of asynchrony in species fluctuation stabilize the community (Loreau et al. 2002, Ruijven and Berendse 2007, Blüthgen et al. 2016, Craven et al. 2018). It has been also shown that stability can associate more strongly with the degree of synchrony among dominant species than with species richness (Blüthgen et al 2016, Valencia et al 2020). [While community temporal stability should be related to the stability of ecosystem functions and services, there is limited evidence linking both processes on multitrophic systems.]()

Since the reproductive success of the most plant highly depends on pollination by animals, the stability of pollinator communities may increase the plant reproductive success. A greater diversity of pollinators has been shown to increase the plant reproductive success as a result of complementary service provision arising from variation across of pollinator effectiveness, including space and time (Hoehn et al. 2008; Fründ et al. 2010, Albrecht et al. 2012, Blüthgen and Klein, 2013, Fründ et al. 2013). Pollinator species may exhibit different ranges of temporal activity (Pisanty et al. 2016, Rader et al. 2016) and consequently may provide pollination services at different times of the season. Other studies suggesting that a few dominant species are sufficient to provide a higher reproductive success of plants, i.e pollinator communities are dominated by a small number of common species, and that abundance fluctuations of these dominant species are often responsible for a majority of pollination service provision across an ecosystem (Kleijn et al 2015, Winfree et al 2015).

[gap]()

In the current study, we use a 5-year dataset in Mediterranean woodland patches to study the role of richness and synchrony of pollinator species as drivers of stability of visitation rate. We also evaluate whether the stability of plant reproductive success, employed the stability of fruit proportion and stability of seed numbers, is affected by the stability of pollinator communities. In addition, we analysis whether the effect of the pollinator richness and visitation rate on the reproductive success of the plants varies according to the year.

@@

Our study was conducted in the southwest of the Iberian Peninsula (Huelva and Seville - Spain). We selected 13 sites in Mediterranean shrub ecosystems that ranged from sites with little to no disturbances to more modified ones with crops and urban areas nearby. These sites had similarity in plant community composition and some of most abundance plant species provide nectar and pollen that attract flower visitors. Each site was surveyed at least 7 times per year during the flowering season from 2015 to 2019 (from March to June). Within each site, we selected 3-4 individuals belonging to 3-9 plant species (mean  ± standard error: 5.92  ± 0.54 plant species per site). All the floral visitors (from now on referred to as pollinators) that contacted with their flowers  were registered in visits per flower and minutes by performing timed focal observations to the selected individuals. Only pollinators that could not be identified in the field were captured, stored, and identified to species level in the laboratory. For each plant individual, we counted the total number of flowers and at the end of the season, we recorded the number of fruit and the seeds per fruit.
For each site and year, we calculated the visitation rate of pollinators as the number of visits divided by the number of flowers, pollinator richness, the fruit set as the proportion of fruit, and the seed set as number of seed for each plant individual sampled per site.

@@

For the study of stability, we used 5 plant species (*Cistus crispus*, *Cistus ladanifer*, *Cistus salviifolius*, *Halimium halimifolium*, and *Lavandula pedunculata*) because of they were recorded in at least 5 sites. The stability of visitation rate was quantified as the inverse of the coefficient of variation (mean/SD) for each plant species in each plot following (Tilman, 1999). To examine the role of richness and synchrony of pollinator species as drivers of stability of visitation rate, we calculate the richness as the total number of pollinator species that visit each plant species in each plot. For synchrony, we used the Loreau and de Mazancourt index (2008), the weighted form of the synchrony index (Gross et al., 2014; Blüthgen et al., 2016), and the log of variance ratio (log VR) (Schluter, 1984; Hallett et al., 2014; Lepš et al., 2018). All of them were positively correlated within each plant species, therefore, data for synchrony assessed by Loreau and de Mazancourt index are presented and included in our further analyses. This synchrony measurement varies from zero, when fluctuations are perfectly asynchronous, to one, when fluctuations are perfectly synchronous. Finally, we express asynchrony as 1-Loreau and de Mazancourt’s index of synchrony.
We also calculated the stability of fruit proportion and of the seed number for each plant species in each plot.

@@

All the analyses of this study were conducted in R ver. 4.0.2. To evaluate the main drivers of visitation rate stability, first we run correlation test between asynchrony and richness. Due to we observed a positive correlation among both variables, linear mixed models were used separately to evaluate the relationships between stability of visitation rate and pollinator richness, and stability of visitation rate and asynchrony of pollinator including site and plant species as random factor. We also asses the relationship between asynchrony and pollinator richness with linear mixed model. For a deeper understanding, we also analyzed the respond of the stability for each plant species separately.

For our aim to study whether the stability of the visitation rate influenced in the stability of the plant reproductive success, we run linear mixed models using as response variables the stability of fruit proportion or stability of seed number and as explanatory variable the visitation rate stability. Site and plant species were included as factor random.

@@

@@

Stability of visitation rate ranged from 0.64 to 9.24 (mean ± SE: 1.928 ± 0.232), whereas pollinator richness ranged from 2 to 20 (8.128 ± 0.686) and the pollinator asynchrony varied from 0 to 0.998 (0.754 ± 0.032).
Our results showed a significant relationship between stability of visitation rate and asynchrony with greater visitation rate stability when pollinator asynchrony increases (Estimate: 2.903; standard error: 1.134) (Fig 1A),
whereas stability of visitation rate was not influenced by the pollinator richness (Estimate: -0.001; standard error: 0.063) (Fig 1B). We also observed a significant and positive relationship between asynchrony and pollinator richness (Estimate: 0.031; standard error: 0.007) (Supplementary material-Fig 1).
When we analysed the plant species separately, we found no significant differences but we observed positive tendency in asynchrony for *H. halimifolium* and *L. pedunculata*. In the case of richness, *C. crispus*, *C. ladanifer* and *C. salviifolius* was a negative tendecy wile *H. halimifolium* and *L. pedunculata* was positive (Supplementary material-Fig 2).

@@

Stability of fruit proportion varied from 2.62 to 120.66 (mean ± SE: 19.806 ± 3.615), whereas the stability of seed number ranged 1.28 to 55.86 (7.018 ± 1.441).
When we analyzed whether the stability of the visitation rate affected the stability of the plant reproductive success, we did not find relationship between this and the stability of the fruit proportion (Estimate: -1.547; std. error: 2.611) (Fig 2A), however, the stability of seed number showed negative tendency with the stability of visitation rate (Estimate: -1.887; std. error: 0.959) (Fig 2B).

@@

When we analysed whether the effect of pollinator richness and visitation rate is different across the years, we found that both variables showed a significant interaction effect with the year on fruit proportion. Although we observed a saturation of the fruit proportion, in year like 2017 is showed that a greater richness and visitation rate increased the fruit proportion (Richness - Estimate: 0.395; std. error: 0.15; Visitation rate - Estimate: 25.238; std. error: 5.67). However in 2018, the relationship between richness and fruit proportion was negative [(Estimate: -0.147; std. error: 0.05)](Los estimate y std error en este caso son en funcion de 2015 ya que el modelo lo pone como intercept. El Anova(car) nos da Pr(>Chisq) y df, ¿incluyo esta información??) (Fig 3).

>>>>

1. Understanding the mechanisms driving community stability and whether this stability should be related to the stability of ecosystem functions is important for predicting the resilience of communities to changes in biodiversity. While community stability can be influenced by species richness and asynchrony of species,  the relation with the stability of ecosystem functioning provided by this community is less explored, and specially for functions involving several trophic levels such as plant pollination. [For J ecol, I will focus on stability of plant reproductive success, and then link that to the theory of community stability, so just the same sentences you usem, but order upside down]()

2. Using a 5-year dataset across 13 Mediterranean woodland patches, we assessed the effect of the richness and asynchrony of pollinators as drivers of the stability of plant's visitation rate and plants reproductive success. [In addition, we analyzed whether the year together with richness and visitation rate affected the proportion of fruits and the number of seeds.](esto pensamos como ponerlo con los results finales)

3. The stability of plant's visitation rate is more affected by the asynchrony of pollinator species abundances than by pollinator richness. shows a negative tendency. As for the stability of the plant reproductive success, we found no relationship between the stability of fruits and the stability of the visitation rate, however a negative relation was observed between stability of seed and the stability of visitation rate. However, when years are analysed we found that in some years/[bad years](), the effect of pollinator richness and visitation rate on fruit proportion becomes important.

@@

The role of biodiversity in ecosystem function has long been of great interest to researchers (Hooper et al 2005, Gagic et al., 2015, Lefcheck et al 2015, Oliver et al., 2015, Hong et al 2021). We have experimental and observational evidence that biodiversity enhances ecosystem functioning (Cardinale et al., 2006, Cardinale et al., 2012, Tilman et al., 2014), but the role of biodiversity for the stability of ecosystem functioning is less explored (but see Griffin et al., 2009, Grman et al., 2010), specially for functions involving several trophic levels such as plant pollination (Ebeling et al., 2008, Winfree and Kremen 2009, Garibaldi et al 2011, [Lázaro et al 2022](wild Olea europaea communities)). Globally, it is estimated that at least 78% of wild plants and 75% of leading crops depend to some extent on pollination by animals (Ollerton et al., 2011; Kleijn et al., 2015). Understanding the mechanisms driving pollinator community stability and how this relates to the stability of ecosystem functions and services delivered (i.e. pollination) is important for predicting changes in both community structure and ecosystem function.

Stability can be defined [as the inverse of the coefficient of variation of ecosystem properties over time](irrelevant now) (McCann, 2000; Tilman, 1999). Among main driving of stability are the diversity and species asynchrony. A higher number of species might provide a wider range of responses to environmental changes, which would result in a higher stability (McCann, 2000, Tilman et al. 2006, Hector et al. 2010, Xu et al. 2021), although non-significant relationships between diversity and stability have also been observed (Hooper et al. 2005, van der Plas 2019). The stability is also influenced by the degree of asynchrony between species. In this sense, higher levels of asynchrony in species fluctuation stabilize the community (Loreau et al. 2002, Ruijven and Berendse 2007, Blüthgen et al. 2016, Craven et al. 2018). It has been also shown that stability can associate more strongly with the degree of synchrony among dominant species than with species richness (Blüthgen et al 2016, Valencia et al 2020). [While community temporal stability should be related to the stability of ecosystem functions and services, there is limited evidence linking both processes on multitrophic systems.](I like this sentence)

Since the reproductive success of most plant species highly depends on pollination by animals (Ollerton et al 2012), the stability of pollinator communities may increase the stability of plant reproductive success. A greater diversity of pollinators has been shown to increase the plant reproductive success as a result of complementary service provision arising from variation across of pollinator effectiveness, including space and time (Hoehn et al. 2008; Fründ et al. 2010, Albrecht et al. 2012, Blüthgen and Klein, 2013, Fründ et al. 2013). Pollinator species may exhibit different ranges of temporal activity (Pisanty et al. 2016, Rader et al. 2016) and consequently may provide pollination services at different times of the season. Other studies suggesting that a few dominant species are sufficient to provide a higher reproductive success of plants, i.e pollinator communities are dominated by a small number of common species, and that abundance fluctuations of these dominant species are often responsible for a majority of pollination service provision across an ecosystem (Kleijn et al 2015, Winfree et al 2015).[not sure this last sentence is supported by the refs, plus, is in crops only]()

[gap](I think something about temporal variation and pollen limitation)

In the current study, we use a 5-year dataset in Mediterranean woodland patches to study the role of richness and synchrony of pollinator species as drivers of stability of visitation rate. We also evaluate whether the stability of plant reproductive success, employed the stability of fruit proportion and stability of seed numbers, is affected by the stability of pollinator communities. In addition, we analyzed whether the effect of the pollinator richness and visitation rate on the reproductive success of the plants varies over time.

@@

Our study was conducted in the southwest of the Iberian Peninsula (Huelva and Seville - Spain). We selected 13 sites in Mediterranean shrub ecosystems that ranged from sites with little to no disturbances to more modified ones with crops and urban areas nearby. These sites had similarity in plant community composition and some of most abundance plant species provide nectar and pollen that attract flower visitors. Each site was surveyed at least 7 times per year during the flowering season from 2015 to 2019 (from March to June). Within each site, we selected 3-4 individuals belonging to 3-9 plant species (mean  ± standard error: 5.92  ± 0.54 plant species per site). All the floral visitors (from now on referred to as pollinators) that contacted with their flowers  were registered in visits per flower and minute by performing timed focal observations to the selected individuals. Only pollinators that could not be identified in the field were captured, stored, and identified to species level in the laboratory. For each plant individual, we counted the total number of flowers and at the end of the season, we recorded the number of fruit and the seeds per fruit. All plants analyzed are at least partially pollinator dependent.
For each site and year, we calculated the visitation rate of pollinators as the number of visits divided by the number of flowers, pollinator richness, the fruit set as the proportion of fruit, and the seed set as number of seed for each plant individual sampled per site.

@@

For the study of stability, we used 5 plant species (*Cistus crispus*, *Cistus ladanifer*, *Cistus salviifolius*, *Halimium halimifolium*, and *Lavandula pedunculata*) because of they were recorded in at least 5 sites [what about years? in all years?](). The stability of visitation rate was quantified as the inverse of the coefficient of variation (mean/SD) for each plant species in each plot following (Tilman, 1999). To examine the role of richness and synchrony of pollinator species as drivers of stability of visitation rate, we calculate the richness as the total number of pollinator species that visit each plant species in each plot. To calculate synchrony, we used the Loreau and de Mazancourt index (2008), the weighted form of the synchrony index (Gross et al., 2014; Blüthgen et al., 2016), and the log of variance ratio (log VR) (Schluter, 1984; Hallett et al., 2014; Lepš et al., 2018). All of them were positively correlated to each other within each plant species [sup mat?](), therefore, data for synchrony assessed by Loreau and de Mazancourt index are presented and included in our further analyses. This synchrony measurement varies from zero, when fluctuations are perfectly asynchronous, to one, when fluctuations are perfectly synchronous. Finally, we express asynchrony as 1-Loreau and de Mazancourt’s index of synchrony.
We also calculated the stability of fruit proportion and of the seed number for each plant species in each plot.

@@

All the analyses of this study were conducted in R ver. 4.0.2. To evaluate the main drivers of visitation rate stability, first we run correlation test between pollinator asynchrony and richness. As we observed a positive correlation among both variables [indicate r](), linear mixed models were used separately to evaluate the relationships between stability of visitation rate and pollinator richness, and stability of visitation rate and asynchrony of pollinator including site and plant species as random factor. [We also asses the relationship between asynchrony and pollinator richness with linear mixed model.](first you said correlation, now is lm?) For a deeper understanding, we also analyzed the response for each plant species separately.

To study whether the stability of the visitation rate influenced in the stability of the plant reproductive success, we run linear mixed models using as response variables the stability of fruit proportion or stability of seed number and as explanatory variable the visitation rate stability. Site and plant species were included as factor random.

@@

All models check with Dharma... cite packages used for plotting, etc...

@@

Stability of visitation rate ranged from 0.64 to 9.24 (mean ± SE: 1.928 ± 0.232), whereas pollinator richness ranged from 2 to 20 (8.128 ± 0.686) and the pollinator asynchrony varied from 0 to 0.998 (0.754 ± 0.032).
Our results showed a significant relationship between stability of visitation rate and asynchrony with greater visitation rate stability when pollinator asynchrony increases (Estimate ± standard error: 2.903 ± 1.134) (Fig 1A),
whereas stability of visitation rate was not influenced by the pollinator richness (-0.001 ± 0.063)[only need to explain what is what the first time]() (Fig 1B). [We also observed a significant and positive relationship between asynchrony and pollinator richness (Estimate: 0.031; standard error: 0.007) (Supplementary material-Fig 1).](decide if this is needed)
When we analysed the plant species separately, we found [no significant](are we reporting p-values?) differences but we observed positive tendency in asynchrony for *H. halimifolium* and *L. pedunculata* [and other species has also positive estmate, right?](). In the case of richness, *C. crispus*, *C. ladanifer* and *C. salviifolius* was a negative tendecy wile *H. halimifolium* and *L. pedunculata* was positive (Supplementary material-Fig 2).

@@

Stability of fruit proportion varied from 2.62 to 120.66 (mean ± SE: 19.806 ± 3.615), whereas the stability of seed number ranged 1.28 to 55.86 (7.018 ± 1.441).
When we analyzed whether the stability of the visitation rate affected the stability of the plant reproductive success, we did not find a relationship between this and the stability of the fruit proportion (Estimate: -1.547; std. error: 2.611) (Fig 2A), however, the stability of seed number showed negative tendency with the stability of visitation rate (Estimate: -1.887; std. error: 0.959) (Fig 2B).

@@

When we analysed whether the effect of pollinator richness and visitation rate is different across the years, we found that both variables showed a significant interaction effect with the year on fruit proportion. Although we observed a saturation of the fruit proportion, in year like 2017 is showed that a greater richness and visitation rate increased the fruit proportion (Richness - Estimate: 0.395; std. error: 0.15; Visitation rate - Estimate: 25.238; std. error: 5.67). However in 2018, the relationship between richness and fruit proportion was negative [(Estimate: -0.147; std. error: 0.05)](Los estimate y std error en este caso son en funcion de 2015 ya que el modelo lo pone como intercept. El Anova(car) nos da Pr(>Chisq) y df, ¿incluyo esta información??) (Fig 3).[Hay que pensar como lo contamos]()
